On the presentation of orthogonal groups over polynomial rings

### 2019-04-30 dl-paper.tex
Finished editing 4.2

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof}
Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we conclude from this that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$. Thus, we have shown that $P_\alpha(*)$ contains $\St(\Phi, A[X], M[X])$.

Now choose $\gamma$ to be arbitrary root orthogonal to $\alpha$. By~\cref{Kgen-strong} we are left to prove that $c_{\pm \gamma}(f, X\xi)$ lies in $P_\alpha(*)$. 
But $c_{\pm \gamma}(f, X\xi) = x_{\pm \gamma}(f) \cdot z_{\pm \gamma}(-f, -X\xi)$ is a product of two generators of type 3.
\begin{comment}
Let us draw the matrix for $z_\alpha(m, X\xi)$:
\[ \left(\begin{array}{cc} 1 & 0 \\ -X\xi & 1 \end{array} \right) \cdot \left(\begin{array}{cc} 1 & m \\ 0 & 1 \end{array} \right) \cdot \left(\begin{array}{cc} 1 & 0 \\ X\xi & 1 \end{array} \right) = 
 \left(\begin{array}{cc} 1 & m \\ -X\xi & 1 - mX\xi \end{array} \right) \cdot \left(\begin{array}{cc} 1 & 0 \\ X\xi & 1 \end{array} \right) = \left(\begin{array}{cc} 1 + mX\xi & m \\ -mX^2\xi^2 & 1 - mX\xi \end{array} \right) \]
Now let us draw the matrix for $c_\alpha(m, \xi)$:
\[ \left(\begin{array}{cc} 1 & m \\ 0 & 1 \end{array} \right) \left(\begin{array}{cc} 1 + mX\xi & -m \\ mX^2\xi^2 & 1 - mX\xi \end{array} \right) = 
 \left(\begin{array}{cc} 1 + mX\xi + m^2X^2\xi^2 & -m^2X\xi \\ mX^2\xi^2 & 1 - mX\xi \end{array} \right) \] 
\end{comment}
\end{proof}

\begin{comment}
Useful consequence: $z_{-\alpha}(X^2f(X), g(X)) \in P_\alpha(0)$. Proof: $z_{-\alpha}(X^2f(X), g(X)) = 
x_{-\beta}(-Xg) \cdot x_{-\alpha-\beta}(\epsilon X) \cdot x_{\alpha+\beta}(\epsilon X^3f^2g^2) \cdot x_{\alpha}(-X^2fg^2) \cdot 
z_{-\beta}(-Xg, -Xf) \cdot x_{\beta}(-X^3f^2g)  \cdot x_{-\alpha}(-X^2f) \cdot z_{-\alpha-\beta}(\epsilon X, -\epsilon Xfg)$,
where $\epsilon = N_{-\alpha-\beta,\beta}$. This is a product of generators of types: 1,1,2,2,1,1,4,1.
\end{comment}

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r] & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\]
\end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator of $P_\alpha(0)$ listed in~Definition~\ref{defP0} still belongs to $P_\alpha(0)$. The assertion is obvious for the generators of type (3) and (4).

@@

Finally, in the case $\alpha + \beta \in \Phi$ (only possible for the generator of type (1)) we have:
\[ z_{\beta}(Xf, g) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2fg ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2fg^2) \cdot z_{\beta}(Xf, g), \]
the latter expression, clearly, is a product of generators of type $4, 1, 1$.
\end{proof}

@@

Let us verify the assertion for the generator of type (4). Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$ and invoke~\eqref{Z5} of~\cref{Zrels} substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$:
\begin{multline} \nonumber
 z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(f a X^2) \cdot x_{-\alpha}(X^2f) \cdot \\ \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon Xf a^2) \cdot x_{\alpha}(- f a^2) \cdot z_{-\beta}(f a, -X),
\end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 1, 1, 1, 2, 2.

@@

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$. \end{corollary}
\begin{proof}
Suppose $g \in P_\alpha(m)$, then $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$ for which we already know that $S_\alpha(a, g_0) \in P(\alpha, 0)$.
We need to verify that $h = S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)$ lies in $P\left(\alpha, \frac{m}{1 + am}\right)$, but from Lemmas~\ref{P0_conj}--\ref{lem:Tulenbayev-formula} and the antisymmetry of Steinberg symbols we obtain that:
\begin{multline} h = {}^{x_\alpha(aX^{-1})}\!g_0 \cdot \left(x_{\alpha}(aX^{-1}) \cdot x_{-\alpha}(mX)\right) \cdot x_{\alpha}\left(-\frac{aX^{-1}}{1+am}\right) \cdot \{X, 1+am\} \cdot h_\alpha^{-1}(1 + am) = \\
= S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \in P_\alpha\left(\frac{m}{1+am}\right).\qedhere \end{multline}
\end{proof}

\begin{lemma} If $\alpha - \beta \in \Phi$ then for all $g \in \St(\Phi, R[X], XM[X])$ holds $S_\alpha(a, g) \in P(\beta, 0)$. \end{lemma}
\begin{proof}
It suffices to verify the assertion for the generators $z_\gamma(Xf, g)$, $\gamma \in \Phi$.
By our assumption $(\alpha, \beta) \geq 0$. 
\end{proof}

>>>>

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof}
Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we obtain that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$ and hence that $P_\alpha(*)$ contains $\St(\Phi, A[X], M[X])$.

Now choose $\gamma$ to be arbitrary root orthogonal to $\alpha$. By~\cref{Kgen-strong} we are left to prove that $c_{\pm \gamma}(f, X\xi)$ lies in $P_\alpha(*)$. 
But $c_{\pm \gamma}(f, X\xi) = x_{\pm \gamma}(f) \cdot z_{\pm \gamma}(-f, -X\xi)$ is a product of two generators of $P_\alpha(0)$ of type 3.
\end{proof}

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r] & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\]
\end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to the generators of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$. The assertion is obvious for the generators of type (3) and (4).

@@

Finally, in the case $\alpha + \beta \in \Phi$ (only possible for the generator of type (1)) we have:
\[ z_{\beta}(Xf, g) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2fg ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2fg^2) \cdot z_{\beta}(Xf, g). \]
It is clear that the latter expression is a product of generators of type $4, 1, 1$.
\end{proof}

@@

Let us verify the assertion for the generator of type (4). Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$. Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = -\alpha - \beta$ into~\eqref{Z5} of~\cref{Zrels} we obtain that
\begin{multline} \nonumber
 z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(f a X^2) \cdot x_{-\alpha}(X^2f) \cdot \\ \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon Xf a^2) \cdot x_{\alpha}(- f a^2) \cdot z_{-\beta}(f a, -X),
\end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 1, 1, 1, 2, 2.

@@

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$. \end{corollary}
\begin{proof}
Suppose $g \in P_\alpha(m)$, then $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$ for which we already know that $S_\alpha(a, g_0) \in P(\alpha, 0)$.
We need to verify that $h = S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)$ lies in $P\left(\alpha, \frac{m}{1 + am}\right)$, but from Lemmas~\ref{P0_conj}--\ref{lem:Tulenbayev-formula} and the antisymmetry of Steinberg symbols we obtain that:
\begin{multline} \nonumber h = {}^{x_\alpha(aX^{-1})}\!g_0 \cdot \left(x_{\alpha}(aX^{-1}) \cdot x_{-\alpha}(mX)\right) \cdot x_{\alpha}\left(-\frac{aX^{-1}}{1+am}\right) \cdot \{X, 1+am\} \cdot h_\alpha^{-1}(1 + am) = \\
= S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \in P_\alpha\left(\frac{m}{1+am}\right).\qedhere \end{multline}
\end{proof}

\begin{lemma} If $\alpha - \beta \in \Phi$ then for all $g \in \St(\Phi, R[X], XM[X])$ holds $S_\alpha(a, g) \in P_\beta(0)$. \end{lemma}
\begin{proof}
It suffices to verify the assertion for the generators $z_\gamma(Xf, g)$, $\gamma \in \Phi$.
By our assumption $(\alpha, \beta) \geq 0$. 
\end{proof}
